```{r}
library(tidyverse)
```

fig. for protein vs fopt when there is no GFP present (cell baseline)

corresponding python notebook: calibrate_single_population_charging.ipynb

Parameters: 
time_limit = 200
time_step = 5
transcript_lens = 1000
cellular_transcript_copy_number = 100
gfp_transcript_copy_number = 10
ribosome_copy_number = 500
total_trna = 2500
ecol_rbs_rate = 100000.0
ribosome_binding_rate = 0.001
trna_charging_rates = [10.0, 50.0, 100.0, 250.0, 500.0, 1000.0, 5000.0]
transcript_names = ["cellularProtein", "GFP"]
trna_composition = (0.7, 0.3)
ecol_mrna_composition = ecol_mrna_composition
gfp_mrna_compositions = [(0.7, 0.3)]
ribosome_speed = 3 
ribosome_footprint = 15

date = "feb-27-2024"

```{r}
df_prot_cal <- read_csv("../processed_data/calibrate_single_population_charging_1e6.csv")
```

```{r}
df_prot_cal %>% filter(species == "cellularProtein") %>% 
  ggplot(aes(x = ecol_comp, y = protein, color = charging_rate, group = charging_rate, alpha = charging_rate != 250.0)) +
  geom_vline(xintercept=0.7, color="grey70", linewidth=2, alpha=0.5, linetype="dashed") +
  geom_line(linewidth = 2) +
  scale_color_distiller(trans = "log10", 
                     palette = "Blues",
                     name = expression(K[charge]),
                     direction=1,
                     breaks = c(10, 100, 1000),
                     labels = c(expression(10^{1}), expression(10^{2}), expression(10^{3}))) +
  scale_x_continuous(name = expression(f[opt])) +
  scale_y_continuous(name = "protein",
                     breaks=c(0, 50000, 100000),
                     labels=c(0, expression(5*"x"*10^{4}), expression(1*"x"*10^{5}))) +
  scale_alpha_manual(values = c(1, 1), guide = FALSE) +
  theme_bw() +
  theme(text = element_text(size = 16),
        legend.position=c(.1,.7),
        legend.margin=margin(c(5,10,15,10)),
        legend.box.background = element_rect(color = "grey"))
ggsave("fig/cellular_baseline_speed_3.png", width = 6, height = 4.6)
```
Supp. figure of grid search for reasonable parameters (kcharge, kspeed) for gfp simulations

corresponding python notebook: tacc-grid_search_speed_vs_charging.ipynb

```{r}
df_grid <- read_csv("../processed_data/speed_vs_charging_grid_search.csv")
```

```{r}
df_grid_form <- df_grid %>% 
  filter(species %in% c("free ribosome", "ATA_charged", "TTT_charged")) %>%
  pivot_wider(names_from = species, values_from = protein, id_cols = c(speed, charging_rate, ecol)) %>%
  mutate(ATA_charged_perc = ATA_charged / (2500 * 0.3),
         TTT_charged_perc = TTT_charged / (2500 * 0.7),
         free_rib_perc = `free ribosome` / 500) %>%
  filter(ecol == 0.6)
#df_grid_form %>% filter(ecol == 0.6)
```

```{r}
base_plot <- function() {
  ggplot() +
  scale_y_discrete(expand = expansion(mult = c(0, 0))) +
  scale_x_discrete(expand = expansion(mult = c(0, 0)),
                   labels = c(0.12, 0.25, 0.5, 1, 2, 4, 8)) +
  ylab(expression(K[charge])) +
  xlab(expression(K[speed])) +
  theme_bw() +
  theme(panel.border = element_blank(),
        legend.position = "bottom")
}

grid_opt <- base_plot() + 
  geom_tile(data=df_grid_form, aes(y = as.factor(charging_rate), x = as.factor(speed), fill = TTT_charged_perc)) +
  scale_fill_continuous_diverging(palette = "Blue-Red", 
                                  mid = 0.75,
                                  name = expression(T[c]^{opt})) +
  geom_point(aes(x=5, y=3))
  
grid_non_opt <- base_plot() + 
  geom_tile(data=df_grid_form, aes(y = as.factor(charging_rate), x = as.factor(speed), fill = ATA_charged_perc)) +
  scale_fill_continuous_diverging(palette = "Blue-Red", 
                                  mid = 0.75,
                                  name = expression(T[c]^{non-opt})) +
  geom_point(aes(x=5, y=3))


grid_rib <- base_plot() + 
  geom_tile(data=df_grid_form, aes(y = as.factor(charging_rate), x = as.factor(speed), fill = free_rib_perc)) +
  scale_fill_continuous_diverging(palette = "Blue-Red", 
                                  mid = 0.15,
                                  name = expression(R[f])) +
  geom_point(aes(x=5, y=3))

```

```{r, fig.width=9, fig.height=3}
cowplot::plot_grid(grid_opt, grid_non_opt, grid_rib, ncol = 3, labels = c("A", "B", "C"))
ggsave("fig/grid_search.png", width = 9, height = 3)
```

Fig. for simulations of codon-deoptimized GFP expression, for different cellular fopt (0.6, 0.8, 1.0)

corresponding python notebook: gfp_experiment_trna_skew_final.ipynb

simulation parameters:
time_limit = 100
time_step = 5
transcript_lens = [1000, 300]
cellular_transcript_copy_number = 100
gfp_transcript_copy_number = 20
ribosome_copy_number = 500
total_trna = 2500
ecol_rbs_rate = 10000000.0
ribosome_binding_rates = [1000000.0, 3000000.0, 10000000.0, 30000000.0, 100000000.0]
trna_charging_rates = [100.0, 300.0]
transcript_names = ["cellularProtein", "GFP"]
trna_compositions = [(0.7, 0.3)]
gfp_mrna_compositions = [(0.9, 0.1), (0.7, 0.3), (0.5, 0.5), (0.3, 0.7), (0.1, 0.9)]
ribosome_speeds = [0.5, 2]
ribosome_footprint = 15

ecol_mrna_compositions = [(0.6, 0.4), (0.8, 0.2), (1.0, 0)]
date = "august-9-2024"


```{r}
#df_skew <- read_csv("../processed_data/gfp_regression_ecol_codon_use_speed_0.5_trna_0.7.csv")
df_skew <- read_csv("../processed_data/gfp_experiment_trna_07_03_speed_2_ecol_6_8_10.csv")
```

```{r, fig.width=8, fig.height=2.5}
gfp_plot <- df_skew %>% filter(charging_rate == 100, ecol %in% c(0.6, 0.8, 1.0), gfp_rbs_foldx < 10) %>%
  ggplot(aes(GFP, cellularProtein_norm, color = factor(codon))) +
    geom_point(size=2) + 
    geom_smooth(method = "lm", se = FALSE, formula = y ~ x + 1, fullrange=TRUE) +
    facet_wrap(~ecol, labeller = label_bquote(f[opt] ^ cell == .(ecol))) +
    xlab("GFP") +
    ylab("cell protein") +
    scale_y_continuous(limits = c(0.6, 1)) +
    scale_x_continuous(breaks= c(0, 2500, 5000, 7500),
    #                   labels = c(expression(2.5*"x"*10^{3}), 
    #                   expression(5*"x"*10^{3}), 
    #                   expression(7.5*"x"*10^{3}), 
    #                   expression(1*"x"*10^{4}))) +
    ) +
    #scale_x_continuous(limits=c(0, 10000)) + scale_y_continuous(limits=c(0.8, 1.0)) +
    scale_color_discrete_sequential(name=expression(f[opt]^GFP), palette = "viridis") +
    theme_bw() +
    theme(panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank())
gfp_plot
```

```{r}
base_plot <- function (data) {
  data %>%
  ggplot(aes(x=codon, y=tot_protein)) +
  #geom_smooth(se = FALSE, color="black") +
  #geom_point() +
  geom_line(linewidth = 1.25) +
  geom_vline(xintercept = 0.7, alpha=0.5, linewidth=1.25, linetype="dashed") +
  scale_x_continuous(name = expression(f[opt]),
                         breaks = c(0, 0.25, 0.5, 0.75, 1),
                         labels = c("0", "0.25", "0.5", "0.75", "1")) +
  scale_y_continuous(name = "protein") +
  theme_classic() +
  theme(axis.title = element_text(size = 12),
            axis.text = element_text(size = 12),
            legend.title = element_text(size=12),
            axis.ticks = element_blank(),
            axis.text.y = element_blank(),
            axis.text.x = element_blank(),
            legend.position = "none"
  )
}
```

```{r}
df_baseline <- read.csv("../processed_data/gfp_experiment_baseline.csv")
df_baseline <- df_baseline %>% mutate(tot_protein = GFP + cellularProtein)
```

```{r}
exp_data <- df_skew %>% 
  filter(charging_rate == 100, gfp_rbs_foldx == 3.0) %>%
  mutate(tot_protein = GFP + cellularProtein) %>%
  mutate(average_codon_use = ((100 * 1000 * ecol) + (20 * 300 * 3 * codon)) / ((100 * 1000) + (20 * 300 * 3)))

p1_data <- exp_data %>% filter(ecol == 0.6)
p1 <- base_plot(df_baseline) + 
  geom_point(data=p1_data, 
             aes(x = average_codon_use, y=tot_protein, color = factor(average_codon_use)),
             size = 2.5) +
  scale_color_discrete_sequential(palette = "viridis") +
  geom_vline(xintercept = 0.6, alpha=0.5, linewidth=1.25)

p2_data <- exp_data %>% filter(ecol == 0.8)
p2 <- base_plot(df_baseline) + 
  geom_point(data=p2_data, 
             aes(x = average_codon_use, y=tot_protein, color = factor(average_codon_use)),
             size = 2.5) +
  scale_color_discrete_sequential(palette = "viridis") +
  geom_vline(xintercept = 0.8, alpha=0.5, linewidth=1.25)


p3_data <- exp_data %>% filter(ecol == 1)
p3 <- base_plot(df_baseline) + 
  geom_point(data=p3_data, 
             aes(x = average_codon_use, y=tot_protein, color = factor(average_codon_use)),
             size = 2.5) +
  scale_color_discrete_sequential(palette = "viridis") +
  geom_vline(xintercept = 1.0, alpha=0.5, linewidth=1.25)

opt_plot <- cowplot::plot_grid(p1, p2, p3, nrow = 1, rel_widths = c(1, 1, 1))
```

```{r}
legend <- get_legend(gfp_plot)
p1 <- cowplot::plot_grid(gfp_plot + theme(legend.position = "none"), opt_plot, ncol = 1, align = 'v', labels = c("A", "B"))
cowplot::plot_grid(p1, legend, rel_widths = c(1, 0.2))
ggsave("fig/gfp_experiment_cell_fopt_6_8_10.png", width = 7, height = 4)
```